### Factor analysis of language input measures

```{r LENA Prep}
awc_t1 <- read_csv("data/lena_averages.csv") %>% 
  rename(Subject = ParticipantID)

names(awc_t1) <- names(awc_t1) %>% 
  str_replace("Prop_", "") %>% 
  str_replace("_Hourly", "")

# Kids with unreliable LENAs
lena_exc <- constants$lena$exclude$timepoint1 %>% map_chr("id")
  
# CI users
ci_users <- constants$exclude %>% map_chr("id")

# Apply exclusions
lena_t1 <- awc_t1 %>%
  filter(constants$lena$rules$min_hours <= Hours,
         Subject %not_in% lena_exc,
         Subject %not_in% ci_users) %>%
  select(-Study, -Hours, -Silence, -Noise, -CVC) %>% 
  rename(WordsPerHour = AWC, CTCPerHour = CTC)

# Count kids
lena_efa <- list(
  n_kids = n_distinct(awc_t1$Subject),
  n_left = n_distinct(lena_t1$Subject)
)

just_stats <- lena_t1 %>% select(-Subject)
```

```{r LENA Percentiles}
# Load norms from helper file
lena_norms <- constants$lena$norms %>%
  # Convert 10 lists into 10 one-row dfs and combine
  lapply(as_data_frame) %>%
  bind_rows

# Compute sample percentiles for 12-Hour AWCs
sample_percentiles <- (lena_t1$WordsPerHour * 12) %>%
  quantile(probs = lena_norms$Percentile / 100) %>%
  round

# Attach to norms
lena_norms <- lena_norms %>%
  mutate(Sample = sample_percentiles) %>%
  arrange(desc(Percentile))

kables$lena_percentiles <- kable(
  lena_norms,
  format = "markdown",
  col.names = c("Percentile", "Norm AWC", "Sample AWC"),
  caption = "Percentiles of 12-hour adult word counts for LENA norms and this sample.",
  align = c("r", "r", "r"))
```

```{r LENA PCA}
# Create a data-frame summary of a PCA
tidy_pca <- function(pca) {
  eigenvalues <- pca$values
  data_frame(Component = seq_along(eigenvalues), Eigenvalue = eigenvalues)
}

pca_lena <- just_stats %>%
  na.omit %>%
  principal(nfactors = 0, rotate = "none") %>%
  tidy_pca

p_scree <- ggplot(pca_lena) +
  aes(x = Component, y = Eigenvalue) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_point(size = 3) +
  geom_line() +
  labs(x = "Component", y = "Eigenvalues of components") +
  theme_bw(base_size = 12)
```

```{r LENA EFA}
# Make a shortcut fa function that also skips models which raise warnings
fa_ml_promax <- function(...) {
  # Store the arguments in a list so we can extract the nfactors argument
  dots <- list(...)
  args <- c(dots, rotate = "promax", fm = "ml", oblique.scores = TRUE)
  # Try the model. If there's a warning, print the number of factors, print the
  # warning, and return NULL instead.
  tryCatch(do.call(fa, args), warning = function(c) {
    warnings <- paste0("Skipping model with nfactors = ", args$nfactors,
                       ": ", conditionMessage(c))
    warning(warnings, call. = FALSE)
    NULL
  })
}

fa_base <- list(
  a = fa_ml_promax(just_stats, nfactors = 1),
  b = fa_ml_promax(just_stats, nfactors = 2),
  c = fa_ml_promax(just_stats, nfactors = 3)
)

```

```{r R Version}
r_vers <- sprintf("%s.%s", R.version$major, R.version$minor)
```

Recordings from `r lena_efa$n_kids` were collected. Of these, 
`r length(lena_exc)` were excluded because the recording was less than 10 hours
long or because of reported difficulty using the recorder. An additional 
`r length(ci_users)` recordings from children with cochlear implants were
excluded.

The children in this sample conformed to prior norms [@LENATalk, p. 10]. Table
1 shows the Adult Word Count percentiles for this sample and for the sample
used to establish the LENA norms.

```{r LENA norms}
kables$lena_percentiles
```

We performed an exploratory factor analysis using the measurements of language
input from the LENA recordings. We used the average hourly adult word count,
average hourly conversational turn count, and average hourly proportion of
meaningful speech, distant speech, and television. We did not include the
hourly proprortions of noise and silence, because these inputs were just the
absense of speech. We expected that the speech-related input measures (adult
words, conversation turns, meaningful speech) would load onto a coherent
language input factor. We had no a priori hypotheses about the other measures
of audio environment, except for the expectation TV would be a negative
indicator for the language input factor because added television input would
take away from other sources of input.

Analyses were performed in `R` (vers. `r r_vers`). Exploratory factor analysis
was conducted using the `psych` package [vers. `r packageVersion("psych")`, 
@psych], and confirmatory factor analysis was performed using `lavaan` [vers. 
`r packageVersion("lavaan")`, @lavaan].

We initially explored the factor structure of the data using principal
components analysis. The eigenvalue decomposition suggested that 2--3 factors
may underlie these five measures. We conducted a series of exploratory factor
analyses, examining models with 2, and 3 underlying factors. All analyses used
promax (oblique) rotation and maximum likelihood estimation. The 2-factor and
3-factor models could not be estimated, due to Heywood cases. Nevertheless, in
the two-factor model, adult word count, conversational turn count, and
meaningful speech measures loaded strongly onto a shared factor. The second
factor contained a Heywood case (the loading from TV).

```{r LENA CFA Mardia}
cfa_data <- lena_t1 %>%
  select(-Subject) %>%
  mutate(WordsPerHour = WordsPerHour / 1000,
         CTCPerHour = CTCPerHour / 100)

# Pre-format the mardia test right now
mard_line <- pretty_mardia(cfa_data, "kurtosis")
```

```{r LENA CFA, fig.width = 5, fig.height = 5}
lena <- "
  input =~ WordsPerHour + Meaningful + CTCPerHour + TV + Distant
  Distant ~~ TV
"
m_lena <- cfa(lena, cfa_data, estimator = "MLR")
# summary(m_lena, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)

# # Bayes
# library("blavaan")
# b_lena <- bcfa(lena, cfa_data)
# summary(b_lena, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)
# semPaths(m_lena)
# plot(b_lena)
```

```{r LENA CFA Extras}
# Get SRMR and r-squareds for reporting
srmr <- fitmeasures(m_lena) %>% getElement("srmr") %>% round(3)
srmr_line <- pretty_eq("SRMR", srmr)

r2 <- inspect(m_lena, "rsquare") %>%
  magrittr::multiply_by(100) %>%
  round(1) %>%
  as.list

# Make a halfway decent table from the parameter estimates
factor_params <- m_lena %>%
  parameterEstimates(ci = FALSE, standardized = FALSE) %>%
  # No intercepts
  filter(op != "~1") %>%
  # Round values and spell out what lavaan operators mean
  mutate_each(funs(fixed_digits(., 3)), est, z) %>%
  mutate(se = fixed_digits(se, 4),
         pvalue = format_pval(pvalue),
         op = expand_ops(lhs, op, rhs)) %>%
  # Reorder and rename columns
  select(op, lhs, rhs, est:pvalue) %>%
  rename(`Estimate` = est, SE = se, `_z_` = z, `_p_` = pvalue, Type = op)

kables$cfa <- kable(
  factor_params, format = "markdown",
  align = c("l", "l", "l", "r", "r", "r", "r"))
```

Based on the findings of the exploratory factor analysis, we conducted a
confirmatory factor analysis to examine how the five LENA statistics measured a
single latent variable for language input. The data were not multivariate normal
(`r mard_line`), so the model was fit using maximum likelihood with robust
standard errors and a scaled test statistic. The disturbances of the TV and
distant speech indicators were allowed to covary because these variables served
as weak or negative indicators of language input. (Put differently, they
competed with the meaningful speech indicator.) To ensure that the indicators
had similar scales of measurement, the adult word counts values were divided by
1000 and the conversational turn counts by 100. The metric of the language
input factor was fixed to the (kilo-) words per hour indicator.

The estimated model is shown in Figure 1 with standardized parameter estimates
to ease interpretation. Parameter estimates are produced in Table 2.

```{r LENA Factor Diagram}
cfa_fig <- semPaths(
  optimizeLatRes = TRUE,
  mar = c(2, 10, 2, 10),
  curve = 2.5,
  edge.label.cex = .9,
  m_lena,
  nCharNodes = 0,
  intercepts = FALSE,
  whatLabels = "std",
  nDigits = 3,
  layout = "tree",
  rotation = 2,
  sizeLat = 10,
  sizeMan = 10,
  nodeLabels = c("Words \nPer Hour", "Meaningful\nSpeech", 
                 "Conv. Turns\nPer Hour", "TV", "Distant\nSpeech", 
                 "Language\nInput"))
```


The model obtained an adequate fit of the data [`r pretty_model_fit(m_lena)`]. 
The input factor explaned very little of the variance in the TV
and distant speech indicators (`r r2$TV`% and `r r2$Distant`% respectively). It
did however account for a large amount of the variance in converational turns
(`r r2$CTCPerHour`%), adult word count (`r r2$WordsPerHour`%), and proportion
of meaningful speech (`r r2$Meaningful`%).

```{r LENA CFA table}
kables$cfa
```
