---
title: "Using Language Input and Lexical Processing to Predict Vocabulary Size"
author: "Tristan Mahr"
date: "November 2015"
output:
  html_document:
    css: assets/custom.css
    fig_caption: yes
    keep_md: no
    theme: flatly
  md_document:
    variant: markdown_github
    preserve_yaml: yes
  word_document:
    reference_docx: assets/tnr_style.docx
csl: assets/apa.csl
bibliography: assets/refs.bib
---

```{r knitr setup, echo = FALSE, message = FALSE, warning = FALSE}
library("knitr")
# working dir for evaluating code (use parent directory)
opts_knit$set(root.dir = "../")

opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  fig.align = "center",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.path = "./assets/figure/")

# Stores formatted tables until time to drop them in text
kables <- list()
```

```{r packages}
# my eye-tracking package github.com/tjmahr/lookr
library("lookr")

# readers
library("readr")
library("yaml")

# data-frame tools
library("dplyr")
library("tidyr")
library("broom")

# other tools
library("stringr")
library("purrr")

# stats
library("psych")
library("lavaan")
library("semPlot")

# output
library("knitr")
library("ggplot2")

source("R/utils.R")
constants <- yaml.load_file("data/constants.yaml")
```

## Some Helpful Context

I originally wrote this report in May 2015 as a final project for a structural 
equation modeling course. That context should explain the methodical, 
showing-my-work level of detail in the analysis sections and the comparatively 
thin introductory and discussion sections. Two weeks after completing the 
project, I submitted an abstract based on it for consideration at the 2015 
Psychonomics conference. I'm including that original report and the scripts and
data-sets used to produce it as a "full text" companion to the poster.

***

## Abstract

Children learn words by listening to caregivers, and the 
quantity and quality of early language input predict later language 
development. Recent research suggests that lexical processing efficiency 
may mediate the relationship between input and vocabulary growth. We 
asked whether language input and lexical processing at 28-39 months 
predicted vocabulary size one year later in 180 preschoolers. Input was 
measured using adult word counts, conversational turn counts, and 
proportion of meaningful speech and television in the child’s 
environment. Lexical processing was measured using accuracy and rate of 
change on a four-image looking-while-listening eyetracking task. 
Structural equation modeling showed that both input and lexical 
processing predicted vocabulary size, but most of the effect of input 
was mediated by lexical processing. Further, amount of television in the 
environment negatively influenced vocabulary size. These results further 
support the hypothesis that children must efficiently process ambient 
language input to capitalize on learning opportunities. 

## Introduction

Children learn language from their environment. In a sense, children are
language-learning machines, converting language input from their caregivers
into linguistic knowledge and spoken language of their own. When deprived of
this input, they cannot learn language. This is why children with cochlear
implants are said to have two ages: a biological age and a hearing age, the
amount of time with access to spoken languge input.

Moreover, we also know that the quantity of input matters for language
acquisition. The amount of early language input from caregivers predicts
language development later on [e.g., @HartRisley; @Huttenlocher1991;
@Hoff2003]. Children who do not receive _enough_ spoken languge input at home
are at a disadvantage when they enter the classroom.

Children who receive more language input have more opportunities for learning 
and therefore learn more. Let's call this idea the **additive model of language 
input**. The additive model is _raison d'être_ for the LENA Research Foundation 
who offers what is essentially a pedometer for spoken language input. A child
wears a small recording over the course of a typical day, and proprietary LENA
software analyzes the recording to generate a summary of the child's language
input and auditory environment. The LENA Research Foundation has initiated
high-profile community interventions (e.g., Providence Talks, the Thirty Million
Words project) built around their device and this additive model of input.

It's clear that the amount of spoken input matters, but it's not entirely clear 
how ambient language input from caregivers translates in linguistic knowledge 
and spoken language. There are indications that young children need to _process_
this spoken language input. Returning the language-learning machine analogy,
children have to do some processing (i.e., mapping the strings of sounds to
familiar objects) on language input in order to capitalize on learning
opportunities. Call this idea the **mediated model of language input**.

Work by Fernald and colleagues has examined lexical processing in young children
with a number of eyetracking studies. @MarchmanFernald2008 found that accuracy
and speed of lexical processing at age two predicted language and working memory
scores at age eight. @Hurtado2008 found that maternal talk at 18 months
predicted lexical processing speed and vocabulary size at 24 months in a group
of Spanish-learning children. Moreover, the effect of input on vocbulary size
was mediated by processing speed. This finding would support a mediated model of
language input, but vocabulary size also mediated the effect of maternal speech
on processing speed. Because processing and vocabulary size were measured at the
same time, we have no way of choosing which mediation path result past explains
the data. At best, we can claim that language input at eighteen months
predicted two closely related language abilities six months later.
@Weisleder2013 essentially replicated this study, finding that processing and
language input at 19 months of age predicted vocabulary size at 24 months.
Lexical processing, however, mediated the effect of input on vocabulary growth,
and because lexical processing temporally preceded the vocabulary measures, we
have a stronger case for the mediated model of language input. The beneficial 
effects of language input on later vocabulary size was determined by the ability
to efficiently process that input.

In the following study, we examine the mediated model of language input by 
analyzing how language input and lexical processing predict later vocabulary 
size on year later. This study examines the same mediational model of
@Weisleder2013 but our study includes older children (many more of them too),
multiple measures of language input, a different language processing task and
measurements, and a direct measure of expressive vocabulary.

## Methods and Measurements

### Participants

Children from English-speaking households were recruited as part of a
longitudinal study. We examined data from the first two time-points of this
study. At Time 1 in the study, the children were 28--40 months of age. During
Time 1 we collected our measures of language input and lexical processing. At
Time 2, we collected the measure of expressive vocabulary when the children
were 40--52 months of age. Children with cochlear implants were recruited for
the longitudinal study but excluded from the present analysis. All children
underwent a hearing screening at both time-points, and they had normal
development per parent report.

### Language input measurements

Language input data was collected using a LENA digital recorder. As noted
earlier, this device is a small recording device that is worn by the child in a
kid-friendly vest. The device records all audible sounds for up to 16 hours.
The recorder and instructions for using it were given to families.
Families were instructed to activate the recorder in the morning and record a
typical day for the child.

After the recording was completed, LENA software analyzed it to generate a
summary of the child's language environment [@LENAAudio;]. The measure it
reports for each recording include:

* Hourly word-counts of adult language in the child’s environment
* Hourly number of child-adult and adult-child conversational turns
* Hourly proportions of meaningful (close) speech, distant speech, noise,
  television/electronics, and silence
* Hourly number of child vocalizations

For each child, we computed the averages of each of these hourly measurements.

### Eyetracking data

As our measure of lexical processing, we used a four-image
looking-while-listening experiment [@FernaldLWL]. In this paradigm, images of
objects are presented onscreen followed by a prompt to view one of the images.
An eyetracker records how the child's gaze location changes in response to
speech. From these data, we can measure the child's _accuracy_ (how often they
fixated on the named image) and _processing speed_ (how quickly they fixated on
the named image).

In our experiment, four photographs of familiar objects were arranged on a
computer display. During a trial, a spoken prompt directed the child to view
one of the images (e.g., _find the fly_). One these images was the target word
(e.g., _fly_). The other three distractor images contained a semantically
related word (_bee_), a phonologically related word (_flag_), and an unrelated
word (_pen_).

A Tobii T60XL eyetracker system recorded the location of a child's gaze on
screen at rate of 60 Hz. Children saw 24 unique trials in an experimental
block. Two blocks of the experiment (each with different trial orderings and
images) were administered.

### Spoken vocabulary (output) measurements

Children also received an expressive vocabulary test [EVT-2, @EVT2]. In this
test, children are presented with images and are asked to name them. For our
analyses, we used growth scale values (a transformation of the raw score into
an age-invariant scale).

## Analyses


```{r, eval = TRUE, child = "01_efa.Rmd"}
```


```{r, eval = TRUE, child = "02_looks.Rmd"}
```


```{r, eval = TRUE, child = "03_mediation.Rmd"}
```

## Discussion

In this study we examined how language input and lexical processing during the 
third year predicted vocabulary size one year later. @Weisleder2013 found that 
lexical processing fully mediated the effect of language input on vocabulary 
growth. We only partially replicated this finding because both the direct and 
indirect effects of language input on vocabulary growth were significant. Given 
that our sample was 10--20 months older than that of @Weisleder2013, it may be
the case the mediating and predictive role of lexical processing diminishes with
age.

### Future Directions

This study is the first draft of a larger exploration of language input,
lexical processing, and vocabulary growth. As we develop this study and
analyses, we will also consider the following directions:

* Inclusion of SES as a predictor of language input.
* Alternative measures of lexical processing such as random slopes and
  intercepts from mixed-effects growth-curve models [e.g., @Mirman2014].
* Supplemental measures of lexical processing from a similar eye-tracking task.
* Examination of whether age or other predictors moderate the mediating role of
  lexical processing.

## Appendix 

### Input Factor Analysis Model Output

```{r}
summary(m_lena, fit.measures = TRUE, rsquare = TRUE)
```

### Latent Growth Curve Model Output

```{r}
summary(lex_proc_fits$f2, fit.measures = TRUE, rsquare = TRUE)
```

### Omnibus Mediation Model Output

```{r}
summary(med_2, fit.measures = TRUE, rsquare = TRUE)
```

## References
